Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex\\ over domains with shellable triangulations

### 2024-07-04 main.tex

Before we outline the main results of this exposition in more details, 
we first review the state of the literature on Poincar\'e--Friedrichs inequalities on specific classes of domains. 
We take the Poincar\'e--Friedrichs inequality for the gradient of scalar functions as our conceptual point of reference.
Given a domain $\Omega \subseteq \bbR^{n}$ and $1 \leq p \leq \infty$,
we search for a constant $C_{\grad,p} > 0$ such that 
\begin{align*}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align*}
Finding a potential of a given gradient $\nabla u$ is generally a nonlinear operation. 
The operator norm of any \emph{linear} potential construction is an upper bound for the Poincar\'e--Friedrichs constant. 
% Even though the existence of such a bounded linear potential construction is not evident a priori, picking the average-free potential 
One example for a bounded linear operator is the average-free potential:
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} % TODO: use same notation as later 
Its operator norm is the constant $C_{\rm P,p} > 0$ in the Poincar\'e inequality 
\begin{align*}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\rm P,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities or Poincar\'e inequalities.

Most research on optimal constants has addressed the case of convex domains. 
% The optimal constants that depend solely on the domain diameter are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, the optimal constants $C_{\grad,p}$ for any $1 \leq p < \infty$ that depends only on the domain diameter is known when the domain is convex.~\cite{bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e inequality~\cite{brenner2008finite,ern2021finite}. 
An open set star-shaped with respect to a internal point is also star-shaped with respect to some open ball within the domain,
implying Poincar\'e inequalities on star-shaped open sets~\cite{hurri1988poincare}.

The situation is less clear cut when we study the curl or divergence operators in vector calculus.
We use the vector field spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \curl \bfu \in \bfL^p(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \divergence \bfu \in L^p(\Omega) \right\}
    .
\end{align*}
Their members are vector fields whose distributional curls and divergences are in Lebesgue spaces. 
This only requires that certain linear combinations of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
The fundamental difference is that the curl and divergence have infinite-dimensional kernels. 
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case.
We are interested in constants for inequalities such as 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\curl,p}
    \| \bff \|_{\bfL^{p}(\Omega)}
    ,
    \qquad 
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\divergence,p}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and whether these constants can be estimated by the operator norms of linear potential operators 
\begin{align*}
    \Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \quad 
    \Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
Not much attention seems to have been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators. 
% even if the domain is convex, 
This sharply contrasts to the study of the gradient operator.

The literature in functional analysis is much scarcer when it comes to any upper bounds for the Poincar\'e--Friedrichs constants. 
We can access some estimates from the literature in the formalism of exterior calculus. 
We are aware of work by Guerini and Savo~\cite{guerini2004eigenvalue}, who estimate the Poincar\'e--Friedrichs constants over convex domains in the Hilbert space setting, when $p=2$. They provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains. This implies upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators.
By duality, that implies lower bounds for the case of full boundary conditions.
Their study also implies sequences of inequalities, such as the following in 3D convex domains:
\begin{align}\label{math:intro:guerinisavo}
    C_{\grad,2} \geq C_{\curl,2} \geq C_{\divergence,2}.
\end{align}
When the domain is star-shaped with respect to a ball, then estimates for the Poincar\'e--Friedrichs constants 
follow from bounds on the operator norms of regularized Poincar\'e operators~\cite{costabel2010bogovskiui}
as mappings between Lebesgue spaces, though these estimates have not been made explicit in the literature yet.
\\

@@

To ensure the correct continuity properties, we require stronger conditions on the triangulations than in the case of gradient potentials. We postulate that the intermediate subtriangulations are manifolds with boundary, and that the aforementioned interface is, more specifically, a boundary submanifold of the intermediate subtriangulation and the new simplex. Under those conditions, which are met by \textit{shellable} triangulations, we define a well-posed auxiliary problem for the curl operator with partial tangential boundary conditions. Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
\\

% TODO: compare with previous paragraphs 
This construction of potentials requires triangulations with special conditions: 
we assemble the triangulation sequentially, adding one simplex after another, such that the interface between the previous intermediate triangulation and the next simplex is always a boundary submanifold of both.  Loosely speaking, this guarantees that the interface with each successive simplex is sufficiently well-behaved. Such triangulations are called \textit{shellable}. Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics~\cite{ziegler2012lectures}, which we only employ in a simple special case. 
In particular, local stars in 2D and 3D triangulations are shellable. 
\\

@@

\begin{remark}
    Our main interest in this manuscript are finite triangulations of compact sets. 
    There is another reason why we insist on the triangulation being finite: 
    if we do not require our simplicial complexes to be finite,
    then a reasonable definition of simplicial complex would need additional topological conditions. 
    % 
    For example, if we modify our definition of simplicial complex to be infinite, 
    then any subset of Euclidean space would give rise to an ``infinite $0$-dimensional complex''. But this obviously does not reflect the topology. 
    %
    A more interesting such example is this:
    the Cantor set $\calC \subseteq [0,1]$ is a compact set 
    that underlies an ``infinite $0$-dimensional simplicial complex''.
    If we place $\calC$ on the x-axis of a 2D coordinate system 
    and connect each member of $\calC$ to $(0,1)$ via a straight line segment, 
    then the resulting set is still compact, even path-connected,  
    but has an ``infinite $1$-dimensional triangulation''.
    % 
    However, none of these infinite families are infinite simplicial complexes in the sense of topology~\cite{spanier1966homology}. 
\end{remark}

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (codimension $1$ members) of this triangulation.

\begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra. 
\end{remark}

When $\calT$ is a triangulation and $T \in \calT$, then $\patch_{\calT}(T)$ denotes the \emph{local patch} or \textit{star} of $T$, 
which the simplicial subcomplex of $\calT$ that contains all supersimplices of $T$ and their subsimplices. Formally,
\begin{gather*}
    \patch_{\calT}(T) := \bigcup_{ T' \in \supsimplex(\calT,T) } \subsimplex(T').
\end{gather*}
A crucial observation is the following.

\begin{lemma}
 Let $\calT$ be an $n$-dimensional simplicial complex and let $S,S' \in \calT$.
 Then either $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \emptyset$ or there exists $S'' \in \calT$
 such that $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \calT_{n}(S'')$ with $S, S' \subseteq S''$.
\end{lemma}
\begin{proof}
 Let $T \in \calT$ be $n$-dimensional.
 We have $T \in \patch_{\calT}(S )$ if and only if all vertices of $S $ are vertices of $T$.
 We have $T \in \patch_{\calT}(S')$ if and only if all vertices of $S'$ are vertices of $T$.
 Consequently, $T \in \patch_{\calT}(S) \cap \patch_{\calT}(S')$ if and only if $T \in \patch_{\calT}(S'')$,
 where $S'' \in \calT$ is the convex closure of $S$ and $S'$.
\end{proof}

We will use a specific notion of connectivity. 
Given an $n$-dimensional simplicial complex $\calT$, 
we call $n$-simplices $S,S' \in \calT$ \emph{strongly connected in $\calT$} if there exists a sequence $S_0=S,S_1,\dots,S'=S_m \in \calT$ such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$. 
We call such a sequence a \emph{strong path} from $S$ to $S'$ in $\calT$. 
Clearly, strongly connected in $\calT$ is an equivalence relation. A \emph{strongly connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ strongly connected if it has only one strongly connected component.

@@

Whenever $T$ is an $n$-dimensional simplex 
\begin{align}\label{math:geometricshapemeasure}
    \kappa(T)
    := 
    \frac{ \diam(T)^{n} }{ n! \vol(T) }
    .
\end{align}
% TODO: Although this is fairly clear upon reading, we could specify in word that $\kappa(T)$ is (equivalent to) the usual shape regularity paremeter, whereas $\rho(\mathcal T)$ is the maximal ratio of volumes between neighboring elements.
Moreover, when $\calT$ is an $n$-dimensional simplicial complex, we introduce 
\begin{align}\label{math:moremeasures}
    \kappa(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \kappa(T),
%     \qquad 
%     \delta(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \frac{ \diam(T) }{ \diam(T') },
    \qquad 
    \rho(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \frac{ \vol(T) }{ \vol(T') }.
\end{align}
Whenever $S,T$ are two $n$-simplices that share a common face $F$ of codimension $1$, we let $\Xi_{S,T} : S \rightarrow T$ denote the affine diffeomorphism 
that preserves $F$. We define 
\begin{align}\label{math:reflectionmeasure}
    \xi(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \| \Xi_{S,T} \|.
\end{align}
This indicator quantifies how much reflection across the shared face distorts the geometry. 
% TODO: I imagine that $\Delta^n$ is a reference simplex, but this has not been introduced earlier.

@@

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, % TODO: what is non-brnaching?
    then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

>>>>

Before we outline the main results of this exposition in more details, 
we first review the state of the literature on Poincar\'e--Friedrichs inequalities on specific classes of domains. 
We take the Poincar\'e--Friedrichs inequality for the gradient of scalar functions as our conceptual point of reference.
Given a domain $\Omega \subseteq \bbR^{n}$ and $1 \leq p \leq \infty$,
we search for a constant $C_{\grad,\Omega,p} > 0$ such that 
\begin{align*}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align*}
Finding a potential of a given gradient $\nabla u$ is generally a nonlinear operation. 
The operator norm of any \emph{linear} potential construction is an upper bound for the Poincar\'e--Friedrichs constant. 
% Even though the existence of such a bounded linear potential construction is not evident a priori, picking the average-free potential 
One example for a bounded linear operator is the average-free potential:
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the constant $C_{\varnothing,\Omega,p} > 0$ in the Poincar\'e inequality 
\begin{align*}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities or Poincar\'e inequalities.

Most research on optimal constants has addressed the case of convex domains. 
% The optimal constants that depend solely on the domain diameter are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, the optimal constants $C_{\grad,\Omega,p}$ for any $1 \leq p < \infty$ that depends only on the domain diameter is known when the domain is convex.~\cite{bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e inequality~\cite{brenner2008finite,ern2021finite}. 
An open set star-shaped with respect to a internal point is also star-shaped with respect to some open ball within the domain,
implying Poincar\'e inequalities on star-shaped open sets~\cite{hurri1988poincare}.

The situation is less clear cut when we study the curl or divergence operators in vector calculus.
We use the vector field spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \curl \bfu \in \bfL^p(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^p(\Omega) : \divergence \bfu \in L^p(\Omega) \right\}
    .
\end{align*}
Their members are vector fields whose distributional curls and divergences are in Lebesgue spaces. 
This only requires that certain linear combinations of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
The fundamental difference is that the curl and divergence have infinite-dimensional kernels. 
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case.
We are interested in constants for inequalities such as 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{\bfL^{p}(\Omega)}
    ,
    \qquad 
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{\bfL^{p}(\Omega)}
    \leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and whether these constants can be estimated by the operator norms of linear potential operators 
\begin{align*}
    \Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \quad 
    \Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
Not much attention seems to have been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators. 
% even if the domain is convex, 
This sharply contrasts to the study of the gradient operator.

The literature in functional analysis is much scarcer when it comes to any upper bounds for the Poincar\'e--Friedrichs constants. 
We can access some estimates from the literature in the formalism of exterior calculus. 
We are aware of work by Guerini and Savo~\cite{guerini2004eigenvalue}, who estimate the Poincar\'e--Friedrichs constants over convex domains in the Hilbert space setting, when $p=2$. They provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains. This implies upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators.
By duality, that implies lower bounds for the case of full boundary conditions.
Their study also implies sequences of inequalities, such as the following in 3D convex domains:
\begin{align}\label{math:intro:guerinisavo}
    C_{\grad,\Omega,2} \geq C_{\curl,\Omega,2} \geq C_{\divergence,\Omega,2}.
\end{align}
When the domain is star-shaped with respect to a ball, then estimates for the Poincar\'e--Friedrichs constants 
follow from bounds on the operator norms of regularized Poincar\'e operators~\cite{costabel2010bogovskiui}
as mappings between Lebesgue spaces, though these estimates have not been made explicit in the literature yet.
\\

@@

To ensure the correct continuity properties, we require stronger conditions on the triangulations than in the case of gradient potentials. We postulate that the intermediate subtriangulations are manifolds with boundary, and that the aforementioned interface is, more specifically, a boundary submanifold of the intermediate subtriangulation and the new simplex. Under those conditions, which are met by so-called \textit{shellable} triangulations, we define a well-posed auxiliary problem for the curl operator with partial tangential boundary conditions. Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
\\

Shellable triangulations are defined by a special constructive property:
we can assemble such triangulations sequentially, adding one simplex after another, such that the interface between the previous intermediate triangulation and the next simplex is always a boundary submanifold of both.
Loosely speaking, this guarantees that the interface with each successive simplex is sufficiently well-behaved.
Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics~\cite{ziegler2012lectures}, which we only employ in a simple special case. 
In particular, local stars in 2D and 3D triangulations are shellable. 
\\

% This construction of potentials requires triangulations with special conditions: 
% we assemble the triangulation sequentially, adding one simplex after another, such that the interface between the previous intermediate triangulation and the next simplex is always a boundary submanifold of both.  Loosely speaking, this guarantees that the interface with each successive simplex is sufficiently well-behaved. Such triangulations are called \textit{shellable}. Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics~\cite{ziegler2012lectures}, which we only employ in a simple special case. 
% In particular, local stars in 2D and 3D triangulations are shellable. 
% \\

@@

\begin{remark}
    Our main interest in this manuscript are finite triangulations of compact sets. 
    There is another reason why we insist on the triangulation being finite: 
    if we do not require our simplicial complexes to be finite,
    then a reasonable definition of simplicial complex would need additional topological conditions. 
    % 
    For example, if we modify our definition of simplicial complex to be infinite, 
    then any subset of Euclidean space would give rise to an ``infinite $0$-dimensional complex''. But this obviously does not reflect the topology. 
    %
    A more interesting such example is this:
    the Cantor set $\calC \subseteq [0,1]$ is a compact set 
    that underlies an ``infinite $0$-dimensional simplicial complex''.
    If we place $\calC$ on the x-axis of a 2D coordinate system 
    and connect each member of $\calC$ to $(0,1) \in \bbR^{2}$ via a straight line segment, 
    then the resulting set is still compact, even path-connected,  
    but has an ``infinite $1$-dimensional triangulation''.
    % 
    However, none of these infinite families are infinite simplicial complexes in the sense of topology~\cite{spanier1966homology}. 
\end{remark}

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
% \begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra. %
% \end{remark}
}

When $\calT$ is a triangulation and $T \in \calT$, then $\patch_{\calT}(T)$ denotes the \emph{local patch} or \textit{local star} of $T$, 
which the simplicial subcomplex of $\calT$ that contains all supersimplices of $T$ and their subsimplices. Formally,
\begin{gather*}
    \patch_{\calT}(T) := \bigcup_{ T' \in \supsimplex(\calT,T) } \subsimplex(T').
\end{gather*}
A crucial observation is the following.

\begin{lemma}
 Let $\calT$ be an $n$-dimensional simplicial complex and let $S,S' \in \calT$.
 Then either $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \emptyset$ or there exists $S'' \in \calT$
 such that $\patch_{\calT}(S) \cap \patch_{\calT}(S') = \patch_{\calT}(S'')$ with $S, S' \subseteq S''$.
\end{lemma}
\begin{proof}
 Let $T \in \calT$ be $n$-dimensional.
 We have $T \in \patch_{\calT}(S )$ if and only if all vertices of $S $ are vertices of $T$.
 We have $T \in \patch_{\calT}(S')$ if and only if all vertices of $S'$ are vertices of $T$.
 Consequently, $T \in \patch_{\calT}(S) \cap \patch_{\calT}(S')$ if and only if $T \in \patch_{\calT}(S'')$,
 where $S'' \in \calT$ is the convex closure of $S$ and $S'$.
\end{proof}

We introduce a specific notion of connectivity. 
Given an $n$-dimensional simplicial complex $\calT$, 
we call $n$-simplices $S,S' \in \calT$ \emph{strongly connected in $\calT$} if there exists a sequence $S_0=S,S_1,\dots,S'=S_m \in \calT$ such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$. 
We call such a sequence a \emph{strong path} from $S$ to $S'$ in $\calT$. 
Clearly, strongly connected in $\calT$ is an equivalence relation. A \emph{strongly connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ \emph{strongly connected} if it has only one strongly connected component.

@@

Whenever $T$ is an $n$-dimensional simplex 
\begin{align}\label{math:geometricshapemeasure}
    \kappa(T)
    := 
    \frac{ \diam(T)^{n} }{ n! \vol(T) }
    .
\end{align}
% TODO: Although this is fairly clear upon reading, we could specify in word that $\kappa(T)$ is (equivalent to) the usual shape regularity parameter, whereas $\rho(\mathcal T)$ is the maximal ratio of volumes between neighboring elements.
Moreover, when $\calT$ is an $n$-dimensional simplicial complex, we introduce 
\begin{align}\label{math:moremeasures}
    \kappa(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \kappa(T),
%     \qquad 
%     \delta(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \frac{ \diam(T) }{ \diam(T') },
    \qquad 
    \rho(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \frac{ \vol(T) }{ \vol(T') }.
\end{align}
Whenever $S,T$ are two $n$-simplices that share a common face $F$ of codimension $1$, we let $\Xi_{S,T} : S \rightarrow T$ denote the affine diffeomorphism 
that preserves $F$. We define 
\begin{align}\label{math:reflectionmeasure}
    \xi(\calT) := \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } \| \Xi_{S,T} \|.
\end{align}
This indicator quantifies how much reflection across the shared face distorts the geometry. 
% TODO: I imagine that $\Delta^n$ is a reference simplex, but this has not been introduced earlier.

@@

\begin{lemma}    
    If an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, % TODO: what is non-branching?
    then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.
